
\section{Background}
\label{sec:background}

Mobile application developers can choose among several existing approaches when
attempting to perform continuous sensing. The availability of some of these
approaches depends on the mobile device's hardware and operating system.

\subsection{Always On}

The basic approach for continuous sensing is to keep the mobile device
\emph{always on}. As an example, on the Android operating system, the sensing
application acquires a wake lock to prevent the device from sleeping and
registers a listener for a specific sensor. Whenever the sensor produces a new
reading, the operating system passes the reading to the application via a
callback function. This operation continues until the listener is unregistered
or until the wake lock is released. The benefit of this approach is that it
achieves the highest accuracy because all the sensor data is delivered to the
application. However, the device cannot be powered down and hence the power
consumption of this approach is high.

\subsection{Duty Cycling}

To avoid keeping the device awake for extended periods of time, \emph{duty
  cycling} allows a device to sleep for fixed, usually regular, periods of
time. An application on Android can implement duty cycling by scheduling a wake
up alarm using the system's alarm service and by ensuring that it has not
acquired a wake lock. When no application has a wake lock, the operating system
can power down the processor. When the wake up alarm fires, the processor wakes
up and notifies the application. Next, the application collects sensor readings
for a short period of time, and then reschedules another alarm.

Duty cycling has several drawbacks. First, events of interest will not be
detected if they occur while the device is sleeping because the application does
not receive any sensor data that were produced at those times. Second, events of
interest may not occur during the short time that the device is awake. These
power transitions are wasteful and expensive (as we show later). The basic duty
cycling approach can be improved by dynamically adjusting the wake up time. When
an application is woken up and detects an event of interest, it continues to
collect sensor readings until events do not occur for some period, after which
it returns to its regular duty cycling schedule. This improves detection
accuracy (e.g., event recall) but at the expense of increased energy
consumption. Unfortunately, finding a good balance between accuracy and power
savings depends on the specific application and the user's behavior. Another
drawback of this approach is that is does not scale well with multiple
applications. It is possible for mutually unaware applications to implement
conflicting wakeup schedules, resulting in little or no sleep.

\subsection{Sensor Data Batching}

An extension to duty cycling is \emph{sensor data batching} in which the device
hardware is able to collect sensor readings while the main processor is
asleep. Upon wake up, the entire batch of sensor data is delivered to the
application(s) that registered a listener for that specific sensor. Unlike duty
cycling, batching requires hardware support (e.g., it is currently only
available on the newly released Google Nexus 5, running Android 4.4), but it
enables applications to receive all the sensor data, and hence provides detect
accuracy similar to the power-on approach. However, batching affects detection
timeliness. Applications may detect events of interest many seconds after they
actually occurred. Also, as with the duty cycling approach, the device may wake
up to find out that no events of interest have occurred in the current batch.

\subsection{Computation Offloading}

Generalizing batching, a peripheral processor can be used to collect and
\emph{compute} on sensor data while the main processor is asleep. The Reflex
system~\cite{reflex} allows developers to offload their own sensor data analysis
algorithms to the peripheral processor. However, developers are exposed to the
full complexity of the architecture, making application development more
difficult. Even simple sensor-driven applications have to be refactored into
distributed programs. Additionally, it is unclear how multiple applications can
program or use the same sensor.

%% The high level of programming complexity is very likely a cause why such an
%% architecture is not supported by current commercial systems.
%% We have the same problem -Ashvin.
% it appears Reflex provides an API to get sensor data, so devs don't need to
% write code for each sensor version

\subsection{Activity Detection}

Given the challenges with full computation offloading, recent mobile devices
limit offloading to a small, predefined set of \emph{activity detection}
algorithms. For example, the Motorola Moto X uses two low-power peripheral
processors to wake-up the device when certain events occur. A dedicated natural
language processor is used to wake up the device when a user says the phrase "OK
Google Now". Also, a contextual processor is used to turn on a part of the
screen when the mobile device is taken out of a pocket, or start the camera
application when the processor detects two wrist twists. Additionally, the
latest Android 4.4 supports specific wake up events based on detecting
"significant motion" or a step, if a device has appropriate sensors. However,
the application developer has no control over the underlying activity detection
algorithms, including customization of parameters. This approach is easy to use,
and it provides good energy savings because the device only wakes up when the
event of interest occurs. However, it only works for applications that can take
advantage of the predefined set of activities.

%% The main advantage of this approach is it ease of use application developers,
%% provided that the platform supports the activities of interest. Also, very
%% good energy savings are achieved as

%% Fundamentally, this approach provides callbacks for pre-defined activities
%% such as walking or specific phrase recognition.

%% This approach works well if the application developer is interested in one of
%% the predefined activities but . However, the approach provides no support for
%% applications that are interested in activities that are not part of the
%% pre-defined set.  this is redundant. -Ashvin

