
\section{\label{sec:Introduction}Introduction}

Today smartphones and tablets are used primarily to run interactive
foreground applications, such as games and web browsers.  As a result,
current mobile devices are optimized for a use case where applications
are used intermittently during the day, in sessions that last for
several minutes.  For example, Alice takes a little break to pick up
her phone, check the weather, read the latest news story and play a
game for a few minutes, and then puts it away to go back to work.  To
maximize battery life, current mobile devices are engineered to go
into sleep state for most of the day when they are not supporting such
interactive usage.

Unfortunately, most mobile platforms are a poor match for a growing
class of mobile applications that perform continuous background
sensing.  Examples range from context-aware applications~\cite{baldauf2007survey,hong2009context}, such as
targeted advertising, and medical applications~\cite{hameed2003application,preuveneers2008mobile,tsai2007usability} that improve our
well-being or even save lives using activity recognition (e.g., fall
detection), to applications that use participatory sensing to get a
better understanding of the physical world, such as noise pollution~\cite{maisonneuve2009citizen,maisonneuve2009noisetube}
monitoring or traffic prediction~\cite{hull2006cartel}.  While the processing demands of
these applications are modest most of the time, they require periodic
collection of sensor readings, which prevents the device from going to
sleep for extended periods of time.  As a result, applications that
perform continuous sensing may cause the device's battery to drain
within several hours.

To improve support for continuous sensing applications the research
community has proposed the use of fully programmable heterogeneous
architectures~\cite{reflex,littlerock,turducken}.  In these
approaches, developers partition their applications to offload the
initial filtering stages of the application to a low-power processor
(or a hierarchy of processors).  When the code running on the low
power processor detects the occurrence of event of interest, it
proceeds to wake up the phone and passes control to the rest of the
application.  While the ability to run custom code on the low-power
hardware provides great flexibility, the significant complexity
inherent in this approach has so far prevented its adoption in
commercial devices.  Instead, smartphone manufacturers, realizing the
potential of sensing applications, have recently incorporated
low-power processors into their architectures, but have limited
application developers to APIs that provide fixed functionality, by
either batching sensor readings or recognizing a small number of
predefined activities that can be used as wake-up
conditions~\cite{android4.4,motox,nexus5}.

In this paper we argue that these APIs are insufficient for supporting
a rich and flexible set of continuous monitoring applications.
Batching is inefficient for applications that depend on infrequent
events and is not appropriate for applications that require crisp
response time.  An activity recognition API provides little
flexibility with no support for applications interested in events that
are not covered by the set of predefined activities.  

Instead, we argue that the sensing API should expose access to
filter-level functionality.  We present {\em Smartsensors}, an
approach that enables programmers to create custom wake-up conditions
by choosing among a set of predefined filtering algorithms and tuning
their parameters.  By providing developer with access to lower-level
filters that implement commonly used algorithms (e.g., FFT,
exponential average), as opposed to higher-level activity detectors,
Smartsensors provide a better balance between flexibility and ease of
deployment.  Smartsensors allows developers to create a broad set of
wake-up conditions that can be used to detect a wide range of
activities.  Simultaneously, they are significantly easier to program
compared to fully programmable offloading.  Moreover, since the
filters are pre-specified, their implementations can be optimized (by
the device manufacturer) for each low-power processor, improving
application portability between devices.  Furthermore, it is possible
to combine the filtering and wake-up operations when applications
register interest in the same set of events.  While Smartsensors could
be programmed directly by individual applications, it is easy to
imagine user-level activity recognition libraries that encapsulate the
functionality of the Smartsensor to provide simple wake-up conditions
for a large number of activities.

To evaluate the benefits of Smartsensors, we built a prototype
implementation that extends a Nexus 4 phone with a low-power sensor
board.  To enable us to conduct controlled and repeatable experiments,
we mounted the Smartsensor prototype on an AIBO ERA 210 robot dog, and
we developed three applications that use accelerometer readings to
detect actions that the robot performs (e.g., walking, standing and
seating).  Experiments conducted on traces collected with this setup,
as well as limited traces collected from human subjects, show that
Smartsensors can reduce the average energy required to run continuous
sensing application by up to 94\%, while matching the detection recall
and precision of an approach that keeps the phone awake at all times.
Moreover, for most of our usage scenarios,the Smartsensor approach
achieves over 95\% of the potential power saving, indicating that an
implementation that supports custom code offloading will
achieve only marginal additional improvements.

The rest of this paper is organized as follows.
Section~\ref{sec:background} discusses existing approaches to
continuous sensing and their limitations.  Section~\ref{sec:approach}
introduces our new approach for continuous monitoring, which lets
developers configure the filters used to implement wake-up conditions.
Section~\ref{sec:prototype} describes our Nexus 4-based prototype.
Sections~\ref{sec:experimentalSetup} and~\ref{sec:results} present our
experimental setup and the results from our evaluation.  Finally,
Sections~\ref{sec:related} and~\ref{sec:conclusion} describe our work
in the context of related work and conclude the paper.

